---
title: "Rumination Time"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

```{r include=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
if (!require("magrittr")) {
  install.packages("magrittr", dependencies = TRUE)
  library(magrittr)
}
#data manipulation
if (!require("dplyr")) {
  install.packages("dplyr", dependencies = TRUE)
  library(dplyr)
}
#drop_na
if (!require("tidyr")) {
  install.packages("tidyr", dependencies = TRUE)
  library(tidyr)
}
#cld
if (!require("multcompView")) {
  install.packages("multcompView", dependencies = TRUE)
  library(multcompView)
}
#%LIKE%
if (!require("data.table")) {
  install.packages("data.table", dependencies = TRUE)
  library(data.table)
}
#glmer
if (!require("lme4")){install.packages("lme4", dependencies = TRUE)
  library(lme4)
}
#lsmeans
if (!require("lsmeans")){install.packages("lsmeans", dependencies = TRUE)
  library(lsmeans)
}

#effects plot
if (!require("effects")) {
  install.packages("effects", dependencies = TRUE)
  library(effects)
}
if (!require("car")) {
  install.packages("car", dependencies = TRUE)
  library(car)
}
if (!require("ggplot2")) {
  install.packages("ggplot2", dependencies = TRUE)
  library(ggplot2)
}

if (!require("ggpubr")) {
  install.packages("ggpubr", dependencies = TRUE)
  library(ggpubr)
}
```

# Read the data

* Filter only the locomotion scores
* Refactor the scores to only 3 classes (1-2 vs 3 vs 4-5)
* Refactor the observation moments

```{r}
load("../Data/AllData.RData")
```

# Data preparation

```{r}
AnalysisData <- AllData %>% filter(
                                SensorType  %like% "ruminatingTimeDataDay"
    ) %>%
   dplyr::mutate(
                  CalvingTime = as.Date(CalvingTime,format = "%Y-%m-%dT%H:%M:%OSZ"),
                  CalvingSeason = case_when(
                    between(month(CalvingTime), 1, 3) ~ "Winter",
                    between(month(CalvingTime), 4, 6) ~ "Spring",
                    between(month(CalvingTime), 7, 9) ~ "Summer",
                    between(month(CalvingTime), 10, 12) ~ "Autumn")
                  ) %>% 
  dplyr::group_by(
    AnimalNumber,
    HerdIdentifier,
    ObservationMoment,
    LocomotionScore,
    ObservationPeriod,
    LactationNumber,
    Parity,
    CalvingSeason
    ) %>% 
  dplyr::summarise(
    SensorValue = mean(SensorValue,na.rm = TRUE),
    SensorValues = n()
    )  %>%  
  dplyr::filter(
    SensorValues == 4
  ) %>%  
  dplyr::arrange(AnimalNumber,HerdIdentifier) %>%
  dplyr::group_by(AnimalNumber, HerdIdentifier) %>%
  dplyr::mutate(LocomotionMoments = length(LocomotionScore)) %>%
  filter(LocomotionMoments == 4) %>%    #4 locomotionscores minimum
  drop_na()

AnalysisData %>% select("HerdIdentifier", "AnimalNumber", "LactationNumber") %>% n_distinct()

hist(AnalysisData$SensorValue,
     main = "Rumination behaviour (min/day)",
     xlab = "Rumination minutes per day")
```


## Baseline model without any variable

```{r}
baselineLMM <- lmer(
                  SensorValue ~ 1 + (1| AnimalNumber), 
                  REML = F,
                  data = AnalysisData
                  )
qqnorm(residuals(baselineLMM))
```

## Full model using the nested repeated measures GLMM

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
LMM <- lmer(
                  SensorValue ~ 
                    LocomotionScore + ObservationPeriod + ObservationMoment + CalvingSeason +
                    ObservationPeriod:ObservationMoment +
                    ObservationPeriod:LocomotionScore +
                    ObservationMoment:LocomotionScore + 
                    ObservationPeriod:LocomotionScore:ObservationMoment + 
                    HerdIdentifier + 
                    Parity +  (1 | AnimalNumber),
                  REML = FALSE,
                  data = AnalysisData
                  )
qqnorm(residuals(LMM))
```


## Comparison of baseline and nested model

```{r}
anova(LMM, baselineLMM, test="Chisq")
```

## Full model

```{r}
LMMdrop <- drop1(LMM, test="Chisq")
if("Pr(>F)" %in% colnames(LMMdrop))
{
  Pvalues <- LMMdrop$`Pr(>F)`
} else 
{
  Pvalues <- LMMdrop$`Pr(Chi)`
}
LMMdrop
```

## Model fit 2-way interactions

```{r}
LMMReduced = update(LMM, . ~ . - LocomotionScore:ObservationPeriod:ObservationMoment)
drop1(LMMReduced, test="Chisq")
```

# Final model

## Model fit

```{r}
LMMReduced = update(LMMReduced, . ~ . - LocomotionScore:ObservationMoment - LocomotionScore:ObservationPeriod)
drop1(LMMReduced, test="Chisq")
```

## Type 3 Analysis of Variance

```{r}
Anova(LMMReduced, ddf="Satterthwaite")
```

## Summary

```{r}
print(summary(LMMReduced, ddf="Satterthwaite"),correlation=FALSE)
```

## Parameter estimate confidence intervals

```{r}
confint(LMMReduced, parm="beta_")
```

## Least square means

```{r}
multcomp::cld(lsmeans::lsmeans(LMMReduced, ~ HerdIdentifier), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(lsmeans::lsmeans(LMMReduced, ~ CalvingSeason), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(lsmeans::lsmeans(LMMReduced, ~ LocomotionScore), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(lsmeans::lsmeans(LMMReduced, ~ ObservationPeriod), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(lsmeans::lsmeans(LMMReduced, ~ ObservationMoment), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(lsmeans::lsmeans(LMMReduced, ~ Parity), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(lsmeans::lsmeans(LMMReduced, ~ ObservationMoment|ObservationPeriod), alpha=0.05, Letters=letters, adjust="tukey")
```

## Interaction term contrasts

```{r}
summary(multcomp::glht(LMMReduced,  lsm(pairwise ~ ObservationMoment|ObservationPeriod, adjust="tukey")))
```

