---
title: "Leg activity"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

```{r include=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
if (!require("magrittr")) {
  install.packages("magrittr", dependencies = TRUE)
  library(magrittr)
}
#data manipulation
if (!require("dplyr")) {
  install.packages("dplyr", dependencies = TRUE)
  library(dplyr)
}
#drop_na
if (!require("tidyr")) {
  install.packages("tidyr", dependencies = TRUE)
  library(tidyr)
}
#cld
if (!require("multcompView")) {
  install.packages("multcompView", dependencies = TRUE)
  library(multcompView)
}
#%LIKE%
if (!require("data.table")) {
  install.packages("data.table", dependencies = TRUE)
  library(data.table)
}
#glmer
if (!require("lme4")){install.packages("lme4", dependencies = TRUE)
  library(lme4)
}
#lsmeans
if (!require("lsmeans")){install.packages("lsmeans", dependencies = TRUE)
  library(lsmeans)
}

#effects plot
if (!require("effects")) {
  install.packages("effects", dependencies = TRUE)
  library(effects)
}
if (!require("car")) {
  install.packages("car", dependencies = TRUE)
  library(car)
}
if (!require("ggplot2")) {
  install.packages("ggplot2", dependencies = TRUE)
  library(ggplot2)
}

if (!require("ggpubr")) {
  install.packages("ggpubr", dependencies = TRUE)
  library(ggpubr)
}
```

# Read the data

* Filter only the locomotion scores
* Refactor the scores to only 3 classes (1-2 vs 3 vs 4-5)
* Refactor the observation moments

```{r include=TRUE}
load("../Data/AllData.RData")
```


# Data preparation

```{r}
AnalysisData <- AllData %>% filter(
                                SensorType%like% "legActivitySummary"  &
                                SensorValue > 100 # Values < 100 very unlikely
    ) %>%
  dplyr::mutate(
                  CalvingTime = as.Date(CalvingTime,format = "%Y-%m-%dT%H:%M:%OSZ"),
                  CalvingSeason = case_when(
                    between(month(CalvingTime), 1, 3) ~ "Winter",
                    between(month(CalvingTime), 4, 6) ~ "Spring",
                    between(month(CalvingTime), 7, 9) ~ "Summer",
                    between(month(CalvingTime), 10, 12) ~ "Autumn")
                  ) %>% 
  dplyr::group_by(
    AnimalNumber,
    HerdIdentifier,
    ObservationMoment,
    LocomotionScore,
    ObservationPeriod,
    LactationNumber,
    Parity,
    CalvingSeason
    ) %>% 
  dplyr::summarise(
    SensorValue = mean(SensorValue,na.rm = TRUE),
    SensorValues = n()
    )  %>%  
  dplyr::filter(
    SensorValues == 4
  ) %>% 
  dplyr::arrange(AnimalNumber,HerdIdentifier) %>%
  dplyr::group_by(AnimalNumber, HerdIdentifier) %>%
  dplyr::mutate(LocomotionMoments = length(LocomotionScore)) %>%
  filter(LocomotionMoments == 4) %>%    #4 locomotionscores minimum
  drop_na()

AnalysisData %>% select("HerdIdentifier", "AnimalNumber", "LactationNumber") %>% n_distinct()


hist(AnalysisData$SensorValue,
     main = "Leg activity (steps/day)",
     xlab = "Leg activity steps per day")
```

# Model building

## Baseline models 

### Linear model

```{r}
baselineLIN <- lmer(
                  SensorValue ~ 1 + (1| AnimalNumber), 
                  REML = F,
                  data = AnalysisData
                  )
plot(baselineLIN)
```

### Non-Linear model

#### Log transformed LMM model

```{r}
baselineLMM <- lmer(
                  log(SensorValue) ~ 1 + (1| AnimalNumber), 
                  REML = FALSE,
                  data = AnalysisData
                  )
qqnorm(residuals(baselineLMM))

```

#### GLMM model

```{r}
baselineGLMM <- glmer(
                  SensorValue ~ 1 + (1 | AnimalNumber), 
                  family=gaussian(link="log"),
                  data = AnalysisData
                  )
qqnorm(residuals(baselineGLMM))
```

Best fit with log transformed model

## Full model using the nested LMM

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
LMM <- lmer(
                  log(SensorValue) ~ 
                    LocomotionScore + ObservationPeriod + ObservationMoment + CalvingSeason + 
                    ObservationPeriod:ObservationMoment +
                    ObservationPeriod:LocomotionScore +
                    ObservationMoment:LocomotionScore + 
                    ObservationPeriod:LocomotionScore:ObservationMoment + 
                    HerdIdentifier + 
                    Parity +  (1 | AnimalNumber),
                  REML = FALSE,
                  data = AnalysisData
                  )
qqnorm(residuals(LMM))
```

## Comparison of baseline and nested model

```{r}
anova(LMM,baselineLMM, test="Chisq")
```

## Full model

```{r}
LMMdrop <- drop1(LMM, test="Chisq")
if("Pr(>F)" %in% colnames(LMMdrop))
{
  Pvalues <- LMMdrop$`Pr(>F)`
} else 
{
  Pvalues <- LMMdrop$`Pr(Chi)`
}
LMMdrop
```

## Model fit 2-way interactions

```{r}
LMMReduced = update(LMM, . ~ . - LocomotionScore:ObservationPeriod:ObservationMoment)
drop1(LMMReduced, test="Chisq")
if (last(Pvalues) < 0.05) {
  finalLMM <- LMM
} else {
  finalLMM <- LMMReduced
}
```

## Model fit without interactions

```{r}
LMMReduced = update(finalLMM, . ~ . - LocomotionScore:ObservationMoment)
```


# Final model

## Model fit

```{r}
drop1(LMMReduced, test="Chisq")
```

## Type 3 Analysis of Variance

```{r}
anova(LMMReduced, ddf="Satterthwaite")
```

## Summary

```{r}
print(summary(LMMReduced, ddf="Satterthwaite"),correlation=FALSE)
```

## Least square means

```{r message=FALSE, warning=FALSE}
multcomp::cld(marginalHerds <- lsmeans::lsmeans(LMMReduced, ~ HerdIdentifier, type = "response"), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(marginalParity <- lsmeans::lsmeans(LMMReduced, ~ Parity, type = "response"), alpha=0.05, Letters=letters, adjust="tukey") 

multcomp::cld(marginalLocomotionScore <- lsmeans::lsmeans(LMMReduced, ~ LocomotionScore, type = "response"), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(marginalLocomotionScore <- lsmeans::lsmeans(LMMReduced, ~ ObservationPeriod, type = "response"), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(marginalLocomotionScore <- lsmeans::lsmeans(LMMReduced, ~ CalvingSeason, type = "response"), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(marginalLocomotionScore <- lsmeans::lsmeans(LMMReduced, ~ LocomotionScore|ObservationPeriod, type = "response"), alpha=0.05, Letters=letters, adjust="tukey")

multcomp::cld(marginalLocomotionScore <- lsmeans::lsmeans(LMMReduced, ~ ObservationPeriod|ObservationMoment, type = "response"), alpha=0.05, Letters=letters, adjust="tukey")

```

## Interaction term contrasts

```{r}
summary(multcomp::glht(LMMReduced,  lsm(pairwise ~ LocomotionScore, adjust="tukey")))
```

## Specific contrasts

```{r}
marginalHerds
Best = c(1,1,0,0,0,0,0,1)
Worst = c(0,0,0,1,1,0,1,0)
contrast(lsmeans::lsmeans(LMMReduced, ~ HerdIdentifier, type = "response"), 
         method = list("Best vs Worst" = Best - Worst))

```

# Interaction plots

## Data preparation

```{r}
# Contrast data for plotting
LSMs<-lsmeans::lsmeans(LMMReduced, pairwise ~ LocomotionScore|ObservationPeriod, type = "response", glhargs=list())
#print(LSMs)
dfPValues <- summary(LSMs)[[2]]
dfLMSs <- summary(LSMs)[[1]]
dfResults <- cbind(dfLMSs[,!(names(dfLMSs) %in% c("df", "ObservationPeriod", "SE"))], dfPValues)
dfPValuesSplit <- dfResults %>% 
                    tidyr::separate(contrast, c("group1", "group2"), " / ", remove=FALSE) %>%
                    #dplyr::arrange(-lsmean) %>%
                    dplyr::mutate(
                                  p.value = round(p.value, 2),
                                  p.value.2 = cut(p.value, 
                                        breaks = c(-Inf,0.001,0.01,0.05,0.1,Inf),
                                        labels = c("***","**","*","†","ns")),
                                  y.position.2 = max(response),
                                  y.position.3 = case_when(contrast == "1-2 / 4-5" ~ y.position.2*1.15, 
                                                           contrast == "1-2 / 3" ~ y.position.2*1.1,
                                                           TRUE ~ y.position.2*1.05)
                    )
```

## Plot

```{r}

ggplot(
  data=transform(dfPValuesSplit, ObservationPeriod = factor(ObservationPeriod, levels= c("Pre Partum", "Post Partum"))), 
  aes(
    x=LocomotionScore, 
    y=response,
    fill=LocomotionScore)) + 
  geom_bar(
    stat="identity"
    ) + 
  coord_cartesian(ylim = c(min(dfPValuesSplit$response)*0.9, 
                           max(dfPValuesSplit$response)*1.15),
                  expand = TRUE) +
  scale_fill_manual(
    values=c("#b3e0a6","#fda50f", "#9e3d22")
    ) +
  geom_text(
    aes(
      label=sprintf("%0.0f", round(response, digits = 0))
      ),
    vjust=4.5, 
    size=3.5
    )+
  geom_errorbar(
    aes(
      ymin=lower.CL, 
      ymax=upper.CL), 
    width=.2,
    position=position_dodge(.9)
    ) +
  facet_wrap(~ObservationPeriod) + 
  labs(
    title = "Number of steps (n/day)", 
    x= "Locomotion score", 
    y=""
    )  + 
  theme_classic() + 
  theme(text=element_text(size=10)) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  stat_pvalue_manual(comparisons = list( c("1-2", "3"), c("3", "4-5"), c("1-2", "4-5")),
                     label.size = 3,
                     data = dfPValuesSplit,
                     y.position =  "y.position.3",
                     label = "p.value.2") + 
  theme(legend.position = "none")
```
