---
title: "Data import for locomotion prediction in SenseOfSensors"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

# Explanation

https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#model-specification

# Data extraction, transformation and loading

## Packages loaded

```{r warning=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
if (!require("readr")) {
  install.packages("readr", dependencies = TRUE)
  library(readr)
  citation("readr")
}
#data manipulation
if (!require("dplyr")) {
  install.packages("dplyr", dependencies = TRUE)
  library(dplyr)
  citation("dplyr")
}
#data manipulation
if (!require("tidyr")) {
  install.packages("tidyr", dependencies = TRUE)
  library(tidyr)
  citation("tidyr")
}
if (!require("stringr")) {
  install.packages("stringr", dependencies = TRUE)
  library(stringr)
  citation("stringr")
}
#%LIKE%
if (!require("data.table")) {
  install.packages("data.table", dependencies = TRUE)
  library(data.table)
  citation("data.table")
}
if (!require("lubridate")) {
  install.packages("lubridate", dependencies = TRUE) 
  library(lubridate)
  citation("lubridate")
}
#random forest
if (!require("caret")) {
  install.packages("caret", dependencies = TRUE)
  library(caret)
}
#Anova
if (!require("car")) {
  install.packages("car", dependencies = TRUE)
  library(car)
}
```

## Data anonymisation

Make sure you have a file with Anonymization.R which contains some salt for the anonymisation

```{r include=TRUE}
source("../Anonymization.R")
#anony
if (!require("anonymizer")) {
  devtools::install_github("paulhendricks/anonymizer")
  library(anonymizer)
}
citation("anonymizer")
anoCols <- c("HerdIdentifier", "AnimalIdentifier", "AnimalEartag")
```

```{r}
LocomotionData <- 
  read_csv(
    "../Data/LocomotionAndBCSData.csv", 
    col_types = cols(
      DaysInMilk = col_integer(), 
      HerdIdentifier = col_character(), 
      LOCO = col_double(), 
      LactationNumber = col_integer())) %>%
  dplyr::filter(LactationNumber > 1,
                LOCO != 3) %>%
  dplyr::mutate(
    CalvingDate = as.Date(CalvingTime, format="%Y-%m-%d"),
    CalvingSeason = case_when(
                    between(month(CalvingDate), 1, 3) ~ "Winter",
                    between(month(CalvingDate), 4, 6) ~ "Spring",
                    between(month(CalvingDate), 7, 9) ~ "Summer",
                    between(month(CalvingDate), 10, 12) ~ "Autumn"),
    AnimalEartag = as.character(AnimalNumber),
    HerdIdentifier = as.character(HerdIdentifier),
    AnimalIdentifier = as.character(AnimalIdentifier),
    LocomotionScore = as.factor(case_when(
                    LOCO <= 2 ~ "1-2",
                    TRUE ~ "3-5")),
    LocomotionScore = relevel(LocomotionScore, ref = "1-2")
    ) %>% # Only select the necesarry columns
  dplyr::mutate_at(vars(anoCols), anonymizer::anonymize, .char = anoSalt) %>%
  dplyr::select(
    AnimalIdentifier,
    HerdIdentifier,
    LactationNumber,
    LocomotionScore,
    ObservationPeriod,
    ObservationDate,
    CalvingSeason
  )

LocomotionDataWeek4 <- LocomotionData %>% dplyr::filter(ObservationPeriod == "WEEK 4")
LocomotionDataWeek8 <- LocomotionData %>% dplyr::filter(ObservationPeriod == "WEEK 8")
```

## BCS data

```{r}
BCSData <- 
  read_csv(
    "../Data/LocomotionAndBCSData.csv", 
    col_types = cols(
      DaysInMilk = col_integer(), 
      HerdIdentifier = col_character(), 
      LactationNumber = col_integer())) %>%
  dplyr::mutate(
    CalvingDate = as.Date(CalvingTime, format="%Y-%m-%d"),
    AnimalEartag = as.character(AnimalNumber),
    HerdIdentifier = as.character(HerdIdentifier),
    AnimalIdentifier = as.character(AnimalIdentifier),
    ObservationPeriod = case_when(
                    ObservationPeriod == "BEGIN DROOGSTAND" ~ "BCSEarlyDryPeriod",
                    ObservationPeriod == "EIND DROOGSTAND" ~ "BCSEndDryPeriod",
                    ObservationPeriod == "WEEK 4" ~ "BCSWeek4",
                    ObservationPeriod == "WEEK 8" ~ "BCSWeek8"),
    Parity = as.factor(case_when(
                    LactationNumber == 2 ~ "2",
                    LactationNumber == 3 ~ "3",
                    TRUE ~ ">3")),
    Parity = relevel(Parity, ref = "2")
    ) %>% # Only select the necesarry columns
  dplyr::mutate_at(vars(anoCols), anonymizer::anonymize, .char = anoSalt) %>%
  dplyr::group_by(
    AnimalIdentifier,
    HerdIdentifier,
    LactationNumber,
    ObservationPeriod,
    Parity
  ) %>%
  dplyr::summarise(BCS = mean(BCS, na.rm = TRUE)) %>%
  tidyr::spread(key = ObservationPeriod, value = BCS, convert = TRUE) %>%
  dplyr::mutate(
    BCSChangeDryPeriod = BCSEndDryPeriod - BCSEarlyDryPeriod,
    BCSChangeTransition =  BCSWeek4 - BCSEndDryPeriod,
    BCSChangePost =  BCSWeek8 - BCSWeek4,
    BCSEndDryPeriodCat = factor(case_when(
      BCSEndDryPeriod < 3.0 ~ "< 3.0",
      between(BCSEndDryPeriod, 3.0, 3.5) ~ "3.0 to 3.5",
      BCSEndDryPeriod > 3.5 ~ "> 3.5"),
      levels= c("3.0 to 3.5", "< 3.0",  "> 3.5")
    ),
    BCSWeek4Cat = factor(case_when(
      BCSWeek4 < 2.5 ~ "< 2.5",
      between(BCSWeek4, 2.5, 2.75) ~ "2.5 to 2.75",
      BCSWeek4 > 2.75 ~ "> 2.75"),
      levels= c("2.5 to 2.75", "< 2.5", "> 2.75")
    ),
    BCSWeek8Cat = factor(case_when(
      BCSWeek8 < 2.25 ~ "< 2.25",
      between(BCSWeek8, 2.25, 2.75) ~ "2.25 to 2.75",
      BCSWeek8 > 2.75 ~ "> 2.75"),
      levels= c("2.25 to 2.75", "< 2.25", "> 2.75")
    ),
    BCSChangeDryPeriodCat = factor(case_when(
      BCSChangeDryPeriod < 0.00 ~ "Decrease",
      between(BCSChangeDryPeriod, 0.0, 0.25) ~ "Increase 0.00 to 0.25",
      BCSChangeDryPeriod > 0.25 ~ "Increase > 0.25"),
      levels= c("Increase 0.00 to 0.25", "Decrease", "Increase > 0.25")
    ),
    BCSChangeTransitionCat = factor(case_when(
      BCSChangeTransition < -0.75 ~ "Decrease > 0.75",
      between(BCSChangeTransition, -0.75, -0.5) ~ "Decrease 0.50 to 0.75",      
      BCSChangeTransition > -0.5 ~ "Decrease < 0.50"),
      levels= c("Decrease 0.50 to 0.75", "Decrease > 0.75", "Decrease < 0.50")
    ),
    BCSChangePostCat = factor(case_when(
      BCSChangePost < -0.25 ~ "Decrease > 0.25",
      between(BCSChangePost, -0.25, 0.00) ~ "Decrease 0.00 to 0.25",      
      BCSChangePost > 0.0 ~ "Increase/Stable"),
      levels= c("Decrease 0.00 to 0.25", "Decrease > 0.25", "Increase/Stable")
    )
  )
quantile(BCSData$BCSWeek4, c(0.33, 0.66), na.rm = TRUE)
quantile(BCSData$BCSWeek8, c(0.33, 0.66), na.rm = TRUE)
quantile(BCSData$BCSChangeTransition, c(0.33, 0.66), na.rm = TRUE)
quantile(BCSData$BCSChangePost, c(0.33, 0.66), na.rm = TRUE)
xtabs(data = BCSData, ~ BCSWeek4Cat + BCSChangeTransitionCat)

quantile(BCSData$BCSEndDryPeriod, c(0.33, 0.66), na.rm = TRUE)
quantile(BCSData$BCSChangeDryPeriod, c(0.33, 0.66), na.rm = TRUE)
xtabs(data = BCSData, ~ BCSEndDryPeriodCat + BCSChangeDryPeriodCat)

xtabs(data = BCSData, ~ BCSWeek4Cat + BCSChangePostCat)
xtabs(data = BCSData, ~ BCSWeek8Cat + BCSChangePostCat)

plot(x= BCSData$BCSEndDryPeriod, y = BCSData$BCSChangeTransition)
write.csv2(BCSData, file="../Data/BCSData.csv", na="", dec = ".", row.names = FALSE)
```

## Combined data

```{r}
Week8Data <- LocomotionDataWeek8 %>%
  inner_join(
    BCSData %>% dplyr::select(
      AnimalIdentifier, 
      HerdIdentifier, 
      LactationNumber,
      Parity,
      BCSEndDryPeriodCat,
      BCSWeek4Cat,
      BCSWeek8Cat,
      BCSChangeDryPeriodCat,
      BCSChangeTransitionCat,
      BCSChangePostCat
    ),
    by = c("AnimalIdentifier", 
           "HerdIdentifier", 
           "LactationNumber")
    ) %>%
  drop_na()

names(Week8Data) <- gsub("[^[:alnum:]]","",names(Week8Data))

xtabs(~Week8Data$BCSWeek4Cat + Week8Data$BCSChangeTransitionCat)
```

# Association model locomotion score week 8

```{r}
if (!require("lme4")) {
  install.packages("lme4", dependencies = TRUE)
  library(lme4)
  citation("lme4")
}
levels(Week8Data$LocomotionScore)
```

## Model Dry and Transition Period


```{r}
Model0.0 <- glmer(formula = LocomotionScore ~ 1 + 
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.0 <- glmer(formula = LocomotionScore ~ Parity +
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.1 <- glmer(formula = LocomotionScore ~ CalvingSeason + 
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.2 <- glmer(formula = LocomotionScore ~ BCSEndDryPeriodCat + 
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.3 <- glmer(formula = LocomotionScore ~ BCSChangeDryPeriodCat +   
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.4 <- glmer(formula = LocomotionScore ~ BCSChangeTransitionCat +  
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.5 <- glmer(formula = LocomotionScore ~ BCSWeek4Cat + 
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.6 <- glmer(formula = LocomotionScore ~ BCSWeek8Cat + 
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.7 <- glmer(formula = LocomotionScore ~ BCSChangePostCat +  
                     (1|HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

```

```{r}
anova(Model0.0, Model1.0)
anova(Model0.0, Model1.1)
anova(Model0.0, Model1.2)
anova(Model0.0, Model1.3)
anova(Model0.0, Model1.4)
anova(Model0.0, Model1.5)
anova(Model0.0, Model1.6)
anova(Model0.0, Model1.7)
```

```{r}
Model2.0 <- glmer(formula = LocomotionScore ~ Parity + CalvingSeason + 
                    BCSEndDryPeriodCat + BCSChangeDryPeriodCat +
                    BCSWeek4Cat + BCSChangeTransitionCat + 
                    BCSWeek8Cat + BCSChangePostCat + 
                    (1 | HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
Model2.1 <- glmer(formula = LocomotionScore ~ Parity + CalvingSeason + 
                    BCSEndDryPeriodCat*BCSChangeDryPeriodCat +
                    (1 | HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
Model2.2 <- glmer(formula = LocomotionScore ~ Parity + CalvingSeason + 
                    BCSWeek4Cat*BCSChangeTransitionCat + 
                    (1 | HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
Model2.3 <- glmer(formula = LocomotionScore ~ Parity + CalvingSeason + 
                    BCSWeek8Cat*BCSChangeTransitionCat + 
                    (1 | HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
Model2.4 <- glmer(formula = LocomotionScore ~ Parity + CalvingSeason + 
                    BCSWeek8Cat + 
                    (1 | HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
Model2.5 <- glmer(formula = LocomotionScore ~ Parity + CalvingSeason + 
                    BCSChangeTransitionCat + 
                    (1 | HerdIdentifier),
                   data = Week8Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

anova(Model2.0, Model2.1, Model2.2, Model2.3, Model2.4, Model2.5, test = "Chisq")
```

```{r}
Anova(Model2.4)
```

# Final model selection

```{r}
finalModel <- Model2.4
```


## Summary

```{r}
summary(finalModel)
Anova(finalModel)
```
## Confidence intervals

```{r message=FALSE, warning=FALSE, eval=FALSE}
#confint(finalModel, parm="beta_")
confint(finalModel)
```

## Least square means and contrasts

```{r message=FALSE, warning=FALSE}
lsmeans::contrast(lsmeans::lsmeans(finalModel, ~Parity, type = "response"), 
          alpha=0.05, 
          method="pairwise", 
          adjust="Tukey")
confint(lsmeans::contrast(lsmeans::lsmeans(finalModel, ~Parity, type = "response"), 
          alpha=0.05, 
          method="pairwise", 
          adjust="Tukey"))
lsmeans::contrast(lsmeans::lsmeans(finalModel, ~CalvingSeason, type = "response"), 
          alpha=0.05, 
          method="pairwise", 
          adjust="Tukey")
confint(lsmeans::contrast(lsmeans::lsmeans(finalModel, ~CalvingSeason, type = "response"), 
          alpha=0.05, 
          method="pairwise", 
          adjust="Tukey"))

lsmeans::contrast(lsmeans::lsmeans(finalModel, ~BCSWeek8Cat, type = "response"), 
          alpha=0.05, 
          method="pairwise", 
          adjust="Tukey")
confint(lsmeans::contrast(lsmeans::lsmeans(finalModel, ~BCSWeek8Cat, type = "response"), 
          alpha=0.05, 
          method="pairwise", 
          adjust="Tukey"))

```