---
title: "Association locomotion score 1-2 vs 3-5 on Week 4 with BCS"
output: 
  github_document:
    toc: true
    toc_depth: 3
editor_options: 
  chunk_output_type: inline
---

# Explanation

https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#model-specification

# Data extraction, transformation and loading

## Packages loaded

```{r warning=FALSE}
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)
if (!require("readr")) {
  install.packages("readr", dependencies = TRUE)
  library(readr)
  citation("readr")
}
if (!require("plyr")) {
  install.packages("plyr", dependencies = TRUE)
  library(plyr)
  citation("plyr")
}
#data manipulation
if (!require("dplyr")) {
  install.packages("dplyr", dependencies = TRUE)
  library(dplyr)
  citation("dplyr")
}
#data manipulation
if (!require("tidyr")) {
  install.packages("tidyr", dependencies = TRUE)
  library(tidyr)
  citation("tidyr")
}
if (!require("stringr")) {
  install.packages("stringr", dependencies = TRUE)
  library(stringr)
  citation("stringr")
}
#%LIKE%
if (!require("data.table")) {
  install.packages("data.table", dependencies = TRUE)
  library(data.table)
  citation("data.table")
}
if (!require("lubridate")) {
  install.packages("lubridate", dependencies = TRUE) 
  library(lubridate)
  citation("lubridate")
}
#random forest
if (!require("caret")) {
  install.packages("caret", dependencies = TRUE)
  library(caret)
  citation(caret)
}
#Anova
if (!require("car")) {
  install.packages("car", dependencies = FALSE)
  library(car)
  citation(car)
}

source(file = "../Functions/mer-utils.R")
```

## Data anonymisation

Make sure you have a file with Anonymization.R which contains some salt for the anonymisation

```{r include=TRUE}
source("../Anonymization.R")
#anony
if (!require("anonymizer")) {
  devtools::install_github("paulhendricks/anonymizer")
  library(anonymizer)
}
citation("anonymizer")
anoCols <- c("HerdIdentifier", "AnimalIdentifier", "AnimalEartag")
```

## Locomotion data

```{r}
LocomotionData <- 
  read_csv(
    "../Data/LocomotionAndBCSData.csv", 
    col_types = cols(
      DaysInMilk = col_integer(), 
      HerdIdentifier = col_character(), 
      LOCO = col_double(), 
      LactationNumber = col_integer())) %>%
  dplyr::filter(LactationNumber > 1) %>%
  dplyr::mutate(
    CalvingDate = as.Date(CalvingTime, format="%Y-%m-%d"),
    CalvingSeason = case_when(
                    between(month(CalvingDate), 1, 3) ~ "Winter",
                    between(month(CalvingDate), 4, 6) ~ "Spring",
                    between(month(CalvingDate), 7, 9) ~ "Summer",
                    between(month(CalvingDate), 10, 12) ~ "Autumn"),
    AnimalEartag = as.character(AnimalNumber),
    HerdIdentifier = as.character(HerdIdentifier),
    AnimalIdentifier = as.character(AnimalIdentifier),
    LocomotionScore = as.factor(case_when(
                    LOCO <= 2 ~ "1-2",
                    TRUE ~ "3-5")),
    LocomotionScore = relevel(LocomotionScore, ref = "1-2")
    ) %>% # Only select the necesarry columns
  dplyr::mutate_at(vars(anoCols), anonymizer::anonymize, .char = anoSalt) %>%
  dplyr::select(
    AnimalIdentifier,
    HerdIdentifier,
    LactationNumber,
    LocomotionScore,
    ObservationPeriod,
    ObservationDate,
    CalvingSeason
  )

LocomotionDataWeek4 <- LocomotionData %>% dplyr::filter(ObservationPeriod == "WEEK 4")
LocomotionDataWeek8 <- LocomotionData %>% dplyr::filter(ObservationPeriod == "WEEK 8")
```

## BCS data

```{r}
BCSData <- 
  read_csv(
    "../Data/LocomotionAndBCSData.csv", 
    col_types = cols(
      DaysInMilk = col_integer(), 
      HerdIdentifier = col_character(), 
      LactationNumber = col_integer())) %>%
  dplyr::filter(LactationNumber > 1) %>%
  dplyr::mutate(
    CalvingDate = as.Date(CalvingTime, format="%Y-%m-%d"),
    AnimalEartag = as.character(AnimalNumber),
    HerdIdentifier = as.character(HerdIdentifier),
    AnimalIdentifier = as.character(AnimalIdentifier),
    ObservationPeriod = case_when(
                    ObservationPeriod == "BEGIN DROOGSTAND" ~ "BCSEarlyDryPeriod",
                    ObservationPeriod == "EIND DROOGSTAND" ~ "BCSEndDryPeriod",
                    ObservationPeriod == "WEEK 4" ~ "BCSWeek4",
                    ObservationPeriod == "WEEK 8" ~ "BCSWeek8"),
    Parity = factor(case_when(
                    LactationNumber == 2 ~ "2",
                    LactationNumber == 3 ~ "3",
                    LactationNumber > 3 ~ ">3")),
    Parity = relevel(Parity, ref = "2")
    ) %>% # Only select the necesarry columns
  dplyr::mutate_at(vars(anoCols), anonymizer::anonymize, .char = anoSalt) %>%
  dplyr::group_by(
    AnimalIdentifier,
    HerdIdentifier,
    LactationNumber,
    ObservationPeriod,
    Parity
  ) %>%
  dplyr::summarise(BCS = mean(BCS, na.rm = TRUE)) %>%
  tidyr::spread(key = ObservationPeriod, value = BCS, convert = TRUE) %>%
  dplyr::mutate(
    BCSChangeDryPeriod = BCSEndDryPeriod - BCSEarlyDryPeriod,
    BCSChangeTransition =  BCSWeek4 - BCSEndDryPeriod,
    BCSEarlyDryPeriodCat = factor(case_when(
      BCSEarlyDryPeriod < 2.75 ~ "< 2.75",
      between(BCSEarlyDryPeriod, 2.75, 3.25) ~ "2.75 to 3.25",
      BCSEarlyDryPeriod > 3.25 ~ "> 3.25"),
      levels= c("2.75 to 3.25", "< 2.75",  "> 3.25")
    ),
    BCSEndDryPeriodCat = factor(case_when(
      BCSEndDryPeriod < 3.0 ~ "< 3.0",
      between(BCSEndDryPeriod, 3.0, 3.5) ~ "3.0 to 3.5",
      BCSEndDryPeriod > 3.5 ~ "> 3.5"),
      levels= c("3.0 to 3.5", "< 3.0",  "> 3.5")
    ),
    BCSWeek4Cat = factor(case_when(
      BCSWeek4 < 2.5 ~ "< 2.5",
      between(BCSWeek4, 2.5, 3.0) ~ "2.5 to 3.0",
      BCSWeek4 > 3.0 ~ "> 3.0"),
      levels= c("2.5 to 3.0", "< 2.5", "> 3.0")
    ),
    BCSChangeDryPeriodCat = factor(case_when(
      BCSChangeDryPeriod < 0.00 ~ "Decrease",
      between(BCSChangeDryPeriod, 0.0, 0.25) ~ "Increase 0.00 to 0.25",
      BCSChangeDryPeriod > 0.25 ~ "Increase > 0.25"),
      levels= c("Increase 0.00 to 0.25", "Decrease", "Increase > 0.25")
    ),
    BCSChangeTransitionCat = factor(case_when(
      BCSChangeTransition < -0.75 ~ "Decrease > 0.75",
      between(BCSChangeTransition, -0.75, -0.5) ~ "Decrease 0.50 to 0.75",      
      BCSChangeTransition > -0.5 ~ "Decrease < 0.50"),
      levels= c("Decrease < 0.50", "Decrease 0.50 to 0.75", "Decrease > 0.75")
    )
  )

print('BCSEarlyDryPeriod')
quantile(BCSData$BCSEarlyDryPeriod, c(0.33, 0.66), na.rm = TRUE)
print('BCSEndDryPeriod')
quantile(BCSData$BCSEndDryPeriod, c(0.33, 0.66), na.rm = TRUE)
print('BCSWeek4')
quantile(BCSData$BCSWeek4, c(0.33, 0.66), na.rm = TRUE)
print('BCSWeek8')
quantile(BCSData$BCSWeek8, c(0.33, 0.66), na.rm = TRUE)
print('BCSChangeDryPeriod')
quantile(BCSData$BCSChangeDryPeriod, c(0.33, 0.66), na.rm = TRUE)
print('BCSChangeTransition')
quantile(BCSData$BCSChangeTransition, c(0.33, 0.66), na.rm = TRUE)

write.csv2(BCSData, file="../Data/BCSDataWeek4.csv", na="", dec = ".", row.names = FALSE)

library("ggpubr")
ggscatter(BCSData, x = "BCSEarlyDryPeriod", y = "BCSChangeDryPeriod", 
          add = "reg.line", conf.int = FALSE, 
          cor.coef = TRUE, cor.method = "pearson")
#BCSEarlyDryPeriodCat*BCSChangeTransitionCat
ggscatter(BCSData, x = "BCSEarlyDryPeriod", y = "BCSChangeTransition", 
          add = "reg.line", conf.int = FALSE, 
          cor.coef = TRUE, cor.method = "pearson")

```


## Dry period data

```{r}
DryPeriodLengthData <- 
  read_csv(
    "../Data/LocomotionAndBCSData.csv", 
    col_types = cols(
      DaysInMilk = col_integer(), 
      HerdIdentifier = col_character(), 
      LactationNumber = col_integer())) %>%
  dplyr::filter(ObservationPeriod == "BEGIN DROOGSTAND") %>%
  dplyr::mutate(
    CalvingDate = as.Date(CalvingTime, format="%Y-%m-%d"),
    AnimalEartag = as.character(AnimalNumber),
    HerdIdentifier = as.character(HerdIdentifier),
    AnimalIdentifier = as.character(AnimalIdentifier),
    ObservationPeriod = case_when(
                    ObservationPeriod == "BEGIN DROOGSTAND" ~ "BCSEarlyDryPeriod",
                    ObservationPeriod == "EIND DROOGSTAND" ~ "BCSEndDryPeriod",
                    ObservationPeriod == "WEEK 4" ~ "BCSWeek4",
                    ObservationPeriod == "WEEK 8" ~ "BCSWeek8"),
    Parity = as.factor(case_when(
                    LactationNumber == 2 ~ "2",
                    LactationNumber == 3 ~ "3",
                    TRUE ~ ">3")),
    Parity = relevel(Parity, ref = "2"),
    DryPeriodLength = as.numeric(difftime(CalvingDate,ObservationDate)),
    DryPeriodLengthCat = factor(case_when(
      DryPeriodLength < 34 ~ "Dry period <34",
      between(DryPeriodLength, 34, 43) ~ "Dry period 34-43",      
      DryPeriodLength > 43 ~ "Dry period >43"),
      levels= c("Dry period 34-43", "Dry period <34", "Dry period >43")
    ),
    DryPeriodLengthCat = relevel(DryPeriodLengthCat, ref = "Dry period 34-43"),
    ) %>% # Only select the necesarry columns
  dplyr::mutate_at(vars(anoCols), anonymizer::anonymize, .char = anoSalt) %>%
  dplyr::filter(between(DryPeriodLength, 0, 1000)) %>%
  dplyr::select(
      AnimalIdentifier, 
      HerdIdentifier, 
      LactationNumber,
      DryPeriodLength,
      DryPeriodLengthCat
  )

quantile(DryPeriodLengthData$DryPeriodLength, c(0.33, 0.66), na.rm = TRUE)
hist(DryPeriodLengthData$DryPeriodLength)

HerdDryPeriodLength <- DryPeriodLengthData %>% group_by(HerdIdentifier) %>% summarise(mean(DryPeriodLength),
                                                                                      quantile(DryPeriodLength, 0.25),
                                                                                      quantile(DryPeriodLength, 0.75))
```

## Combined data

```{r}
Week4Data <- LocomotionDataWeek4 %>%
  inner_join(
    BCSData %>% dplyr::select(
      AnimalIdentifier, 
      HerdIdentifier, 
      LactationNumber,
      Parity,
      BCSEarlyDryPeriodCat,
      BCSEndDryPeriodCat,
      BCSWeek4Cat,
      BCSChangeDryPeriodCat,
      BCSChangeTransitionCat
      
    ),
    by = c("AnimalIdentifier", 
           "HerdIdentifier", 
           "LactationNumber")
    ) %>%
  inner_join(
    DryPeriodLengthData,
    by = c("AnimalIdentifier", 
           "HerdIdentifier", 
           "LactationNumber")
    ) %>%  drop_na() %>% 
  ungroup()

names(Week4Data) <- gsub("[^[:alnum:]]","",names(Week4Data))
```

# Association model locomotion score week 4

```{r}
if (!require("lme4")) {
  install.packages("lme4", dependencies = TRUE)
  library(lme4)
  citation("lme4")
}
```

## Base models

```{r}
Model0.0 <- glmer(formula = LocomotionScore ~ 1 + 
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
```

## Univariate models

```{r}

Model1.0 <- glmer(formula = LocomotionScore ~ Parity +
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.1 <- glmer(formula = LocomotionScore ~ CalvingSeason + 
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.2 <- glmer(formula = LocomotionScore ~ BCSEndDryPeriodCat + 
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.3 <- glmer(formula = LocomotionScore ~ BCSChangeDryPeriodCat +   
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.4 <- glmer(formula = LocomotionScore ~ BCSChangeTransitionCat +  
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.5 <- glmer(formula = LocomotionScore ~ DryPeriodLengthCat +  
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.6 <- glmer(formula = LocomotionScore ~ DryPeriodLength +  
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))


Model1.7 <- glmer(formula = LocomotionScore ~ BCSEarlyDryPeriodCat +  
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))

Model1.8 <- glmer(formula = LocomotionScore ~ HerdIdentifier +  
                     (1|HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))


```

## Univariate comparison

```{r}
anova(Model0.0, Model1.0)
anova(Model0.0, Model1.1)
anova(Model0.0, Model1.2)
anova(Model0.0, Model1.3)
anova(Model0.0, Model1.4)
anova(Model0.0, Model1.5)
anova(Model0.0, Model1.6)
anova(Model0.0, Model1.7)
anova(Model0.0, Model1.8)
```

## Multivariate models

All univariate with P < 0.1

```{r}
Model2.1 <- glmer(formula = LocomotionScore ~ HerdIdentifier + Parity + CalvingSeason + DryPeriodLengthCat + 
                    BCSEarlyDryPeriodCat + BCSChangeTransitionCat + (1 | HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
drop1(Model2.1, test = "Chisq")
print("Variance inflation model 2.1")
vif.mer(Model2.1)
```

## Multivariate models with interactions

```{r}
Model2.2 <- glmer(formula = LocomotionScore ~ Parity + CalvingSeason + DryPeriodLengthCat +
                    DryPeriodLengthCat*BCSEarlyDryPeriodCat + 
                    DryPeriodLengthCat*BCSChangeTransitionCat + 
                    BCSEarlyDryPeriodCat*BCSChangeTransitionCat +
                    (1 | HerdIdentifier),
                   data = Week4Data,
                   family = binomial(link = logit),
                   control=glmerControl(optimizer="bobyqa",
                                        optCtrl=list(maxfun=2e5)))
drop1(Model2.2, test = "Chisq")

print("Variance inflation model 2.2")
vif.mer(Model2.2)
```

# Final model selection

```{r}
finalModel <- update(Model2.1, . ~ . - DryPeriodLengthCat - BCSEarlyDryPeriodCat)
drop1(finalModel, test = "Chisq")
```

A VIF of 1 indicates there is no detectable linear relationship among the regressors, while a large VIF indicates a tight linear relationship. Variance inflation factor, aka VIF. Values over 5 are troubling. We should investigate anything over 2.5.

```{r}
print("Variance inflation final model")
vif.mer(finalModel)
```

## Summary

```{r}
summary(finalModel)
Anova(finalModel)
```

## Confidence intervals

```{r message=FALSE, warning=FALSE, eval=FALSE}
#confint(finalModel, parm="beta_")
confint(finalModel)
```

## Least square means and contrasts

```{r message=FALSE, warning=FALSE}

lsmeans::contrast(lsmeans::lsmeans(finalModel, ~Parity, type = "response"), 
          alpha=0.05, 
          method="revpairwise", 
          adjust="Tukey")
confint(lsmeans::contrast(lsmeans::lsmeans(finalModel, ~Parity, type = "response"), 
          alpha=0.05, 
          method="revpairwise", 
          adjust="Tukey"))
lsmeans::contrast(lsmeans::lsmeans(finalModel, ~CalvingSeason, type = "response"), 
          alpha=0.05, 
          method="revpairwise", 
          adjust="Tukey")
confint(lsmeans::contrast(lsmeans::lsmeans(finalModel, ~CalvingSeason, type = "response"), 
          alpha=0.05, 
          method="revpairwise", 
          adjust="Tukey"))
lsmeans::contrast(lsmeans::lsmeans(finalModel, ~BCSChangeTransitionCat, type = "response"), 
          alpha=0.05, 
          method="revpairwise", 
          adjust="Tukey")
confint(lsmeans::contrast(lsmeans::lsmeans(finalModel, ~BCSChangeTransitionCat, type = "response"), 
          alpha=0.05, 
          method="revpairwise", 
          adjust="Tukey"))
```
